---
title: "sandhi_tri_yang"
author: "Yibing Shi"
date: "2024-12-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggtext)
library(grid)
library(plotly)
library(kml)
```

# Data preparations  
```{r data-construction, cache=FALSE}
load("XSTSF_production.RData")
source('functions.R')
```

datasets
```{r, cache=FALSE}
f0_all_ct <- f0_all_pre %>% filter(focus_condition == 'ct' ) %>% 
  group_by(speaker) %>% 
  mutate(norm_f0 = scale(log(f0))) %>% 
  ungroup() %>% 
  mutate(
    time = as.numeric(time),
    syllable_no = case_when(
    time > 0 & time < 11 ~ 1,
    time > 10 & time < 21 ~ 2,
    time > 20 & time < 31 ~ 3
  ),
  sync_tone1 = ifelse(sync_tone1 == 'RF', 'LHL', sync_tone1),
  sync_tone2 = ifelse(sync_tone2 == 'RF', 'LHL', sync_tone2),
  sync_tone3 = ifelse(sync_tone3 == 'RF', 'LHL', sync_tone3))

tone2_colors  <- c("Ia" = "#E07B91", "Ib" = "#8E735B", "IIa" = "#6B8E66")

f0_tri_ct <- f0_all_ct %>% 
  filter(diortri == 'tri' & sandhi_tone != 'outlier')  %>% 
  mutate(sync_tone23 = paste0(sync_tone2, '_', sync_tone3),
         hist_tone23 = paste0(hist_tone2, '_', hist_tone3),
         hist_tone23_mapped = gsub("yinping", "Ia", 
                         gsub("yangping", "Ib", 
                         gsub("yinshang", "IIa", 
                         gsub("yangshang", "IIb", 
                         gsub("yinqu", "IIIa", 
                         gsub("yangqu", "IIIb", hist_tone23)))))),
         all_tone = paste(sync_tone1, hist_tone23_mapped, sep = "_"),
         tone2_mapped = str_split_fixed(hist_tone23_mapped, "_", 2)[, 1],
         tone23_colored = paste0("<span style='color:", 
                                tone2_colors[f0_tri_ct$tone2_mapped], ";'>",
                                f0_tri_ct$tone2_mapped,  "</span>_<span>", 
                                str_split_fixed(f0_tri_ct$hist_tone23_mapped, "_", 2)[, 2],
                                "</span>"),
         sandhi_tone = ifelse(sandhi_tone == 'HHL', 'MML', 
                              ifelse(sandhi_tone == 'LHL', 'MHL', sandhi_tone)))
  
f0_tri_ct_yp <- f0_tri_ct %>% filter(hist_tone1 == 'yinping') 
f0_tri_ct_yap <- f0_tri_ct %>% filter(hist_tone1 == 'yangping')
f0_tri_ct_ys <- f0_tri_ct %>% 
  filter(hist_tone1 == 'yinshang' & ! ind_no %in% c('S6_3_ct', 'S7_3_ct')) 
f0_tri_ct_yas <- f0_tri_ct %>% filter(hist_tone1 == 'yangshang')
```

## monosyllabic tones
```{r message=FALSE,warning=FALSE}
p1 <- f0_tri_ct_yas %>%
  distri_prop(hist_tone1, sync_tone1, syntax)+
  theme(axis.text.x = element_blank())+
  theme(plot.margin = margin(b = 20),
        legend.title = element_blank())+
  scale_fill_manual(values = c('#C77CFF', '#7CAE00'))+
  labs(fill = "synchronic tone",
       title = 'σ1 (yangshang)');p1

p2 <- f0_tri_ct_yas %>%
  filter(startsWith(hist_tone2, "Ya")) %>%
  distri_prop(hist_tone2, sync_tone2, syntax)+
   theme(axis.text.x = element_blank())+
  theme(plot.margin = margin(b = 20),
        legend.title = element_blank())+
  scale_fill_manual(values = c( '#7CAE00'))+
  labs(fill = "synchronic tone",
       title = 'σ2 (Low-register)');p2
  
p3 <- f0_tri_ct_yas %>%
  filter(startsWith(hist_tone2, "Yi")) %>%
  distri_prop(hist_tone2, sync_tone2, syntax)+
  theme(axis.text.x = element_blank())+
  theme(plot.margin = margin(b = 20),
        legend.title = element_blank())+
  labs(fill = "synchronic tone",
       title = 'σ2 (High-register)');p3

p4 <- f0_tri_ct_yas %>%
  filter(startsWith(hist_tone3, "Ya")) %>%
  distri_prop(hist_tone3, sync_tone3, syntax)+
  theme(axis.text.x = element_blank())+
  theme(plot.margin = margin(b = 20),
        legend.title = element_blank())+
  scale_fill_manual(values = c('#C77CFF', '#7CAE00'))+
  labs(fill = "synchronic tone",
       title = 'σ3 (Low-register)');p4

p5 <- f0_tri_ct_yas %>%
  filter(startsWith(hist_tone3, "Yi")) %>%
  distri_prop(hist_tone3, sync_tone3, syntax)+
  theme(axis.text.x = element_blank())+
  theme(plot.margin = margin(b = 20),
        legend.title = element_blank())+
  labs(fill = "synchronic tone",
       title = 'σ3 (High-register)');p5
```

combination
```{r warning=FALSE}
p <- ggarrange(
  ggarrange(p1 + rremove("ylab") + rremove("xlab"), NULL, 
            ncol = 2, widths = c(1, 1), 
            labels = c('A', ''),
            font.label = list(size = 25, family = "Times")),
  ggarrange(p2 + rremove("ylab") + rremove("xlab"), 
            p3 + rremove("ylab") + rremove("xlab"), 
            ncol = 2, widths = c(1, 1), 
            labels = c('B-1', 'B-2'),
            font.label = list(size = 25, family = "Times")),
  ggarrange(p4 + rremove("ylab") + rremove("xlab"), 
            p5 + rremove("ylab") + rremove("xlab"), 
            ncol = 2, widths = c(1, 1), 
            labels = c('C-1', 'C-2'),
            font.label = list(size = 25, family = "Times")),
  nrow = 3, align = "v", heights = c(1.2, 1.2, 1.2),
  common.legend = TRUE, legend = "top"
); p


p_annotated <- annotate_figure(
  p,
  bottom = text_grob("Historical tonal category", size = 20, family = "Times")
); p_annotated
     

ggsave(filename = file.path(plot_dir, "yash_mono.jpeg"),width = 10, height = 12)
```


# Sandhi categorisation

Auditory categorisation
```{r message=FALSE, warning=FALSE}
unique(f0_tri_ct_yas$sandhi_tone)

p_human <-p_cluster(f0_tri_ct_yas, sandhi_tone)+
  scale_color_manual(values = c("#DDCC77", "#CC6677", "#9B72C7"))+
  theme(legend.position = 'top',
        legend.title = element_blank());p_human
```

*k*-means clustering
```{r}
# data preparation
f0_tri_ct_yas_kmeans <- f0_tri_ct_yas %>% 
  select(-diortri, -syllable_no, -focus_no, -f0) %>% 
  spread(time, norm_f0)

# k-means clustering
cluster_model <- k_means_clustering(f0_tri_ct_yas_kmeans)
kml(cluster_model, nbClusters = 2:10) 
plotAllCriterion(cluster_model)
kml::plot(cluster_model, 3, parTraj=parTRAJ(col="clusters"))


# get cluster results
f0_tri_ct_yas_kmeans <- f0_tri_ct_yas_kmeans %>% 
  mutate(cluster3 = getClusters(cluster_model, 3))

f0_tri_ct_yas_kmeans <- wide_to_long(f0_tri_ct_yas_kmeans) %>% 
  mutate(syllable_no = ifelse(time < 11, '1', 
                              ifelse(time < 21, '2', '3'))) 
```


*k*-means cluster visualisation
```{r warning=FALSE, message=FALSE}
f0_tri_ct_yas_kmeans <- f0_tri_ct_yas_kmeans%>% 
  mutate(cluster3_reorder = case_when(cluster3 == 'B' ~ 'A',
                                      cluster3 == 'A' ~ 'B',
                                      cluster3 == 'C' ~ 'C'))

p_kmeans3 <- p_cluster(f0_tri_ct_yas_kmeans, cluster3)+
  scale_color_manual(values = c("#DDCC77", "#CC6677", "#9B72C7"))+
  theme(legend.position = 'top',
        legend.title = element_blank());p_kmeans3
```

heatmap distribution
```{r message=FALSE, warning=FALSE}
heatmap_df <- heatmap_data(f0_tri_ct_yas_kmeans, cluster3) 
p_htmap3 <- compare_cluster(heatmap_df, 'cluster3'); p_htmap3
```

combination

```{r warning=FALSE}
p <- ggarrange(
  p_human+ labs(title = 'Perceptual clusters'), 
  p_kmeans3+ rremove("ylab") +labs(title = 'k-means clusters \n(k=3)'), 
  p_htmap3 + rremove("ylab") + rremove("xlab")+labs(title = 'Perceptual vs. K-means \n(k=3)'),
  nrow = 1, heights = c(1, 1, 1),
  labels = c('A', 'B', 'C'),
  font.label = list(size = 25, family = "Times",hjust = -1),
  align = "hv",
  legend = "bottom" 
);p


ggsave(filename = file.path(plot_dir, "yash_clusters.jpeg"),width = 12, height = 4)
```


# Comparisons with disyllabic sandhi

plotting sandhi patterns
```{r message=FALSE, warning=FALSE}
# yinshang-initial disyllabic dataset
f0_di_ct_yas <- f0_all_ct %>% 
  filter(syntax %in% c('L', 'M') & diortri == 'di' & hist_tone1 == 'yangshang') %>% 
  filter(is.na(sandhi_tone) == FALSE) 

unique(f0_di_ct_yas$sandhi_tone) # check sandhi categories

# plotting
p_di <- p_cluster_cont(f0_di_ct_yas, sandhi_tone)+
  theme(legend.title = element_blank())+
  scale_color_manual(values = c( "#9B72C7", "#DDCC77"))+
  labs(title = 'Disyllabic sandhi');p_di
p_tri <- p_cluster_cont(f0_tri_ct_yas, sandhi_tone)+
  theme(legend.title = element_blank())+
  scale_color_manual(values = c("#DDCC77", "#CC6677", "#9B72C7"))+
  labs(title = 'Trisyllabic sandhi');p_tri

# check individual contours
# ggplotly(draw_by(f0_di_ct_lcmh_h, 'sandhi_tone'), tooltip = c('text', 'x'))
```


direct comparions between di & tri sandhi
```{r message=FALSE, warning=FALSE}
unique(f0_tri_ct_yas$sandhi_tone)

f0_tri_yas_comp <- f0_tri_ct_yas %>% 
  select(-sync_tone23, -hist_tone23, -hist_tone23_mapped, -tone2_mapped, -tone23_colored, -all_tone) %>% 
  mutate(sandhi_category = case_when(sandhi_tone == 'LHL' ~ 'Rise-fall',
                                     sandhi_tone == 'LMH' ~ 'L-to-H Level',
                                     .default = sandhi_tone),
         sandhi_tone = paste0('tri: ', sandhi_tone))

# direct comparisons of similar sandhi contours
f0_di_yas_comp <- f0_di_ct_yas %>% 
  mutate(sandhi_category = case_when(sandhi_tone == 'LLHH' ~ 'L-to-H Level',
                                     sandhi_tone == 'LLHL' ~ 'Rise-fall',
                                     .default = 'Rise'),
         sandhi_tone = paste0('di: ', sandhi_tone))

f0_yas_comp <- rbind(f0_tri_yas_comp %>% 
                      filter(sandhi_tone %in% c('tri: LMH', 'tri: LHL')),
                    f0_di_yas_comp) %>% 
  mutate(sandhi_tone = factor(sandhi_tone, 
                              levels = c('di: LLHH', 'tri: LMH',
                                'di: LLHL', 'tri: LHL')))
  
p_comp <- p_cluster_cont(f0_yas_comp, sandhi_tone, 'sandhi_category')+
  theme(legend.title = element_blank(),
        legend.position = 'bottom')+
  scale_color_manual(values = c("#CC99CC", "#9B72C7", "#DDCC77", "#AA9944"))+
  labs(title = 'Comparisons of di- and tri-syllabic sandhi');p_comp

unique(f0_di_ct_yas$sandhi_tone)
```

overall distribution comparison
```{r message=FALSE, warning=FALSE}
f0_yas_comp1 <- rbind(f0_tri_yas_comp,f0_di_yas_comp) %>% 
  mutate(sandhi_tone = factor(sandhi_tone, 
                              levels = c('di: LLHL', 'tri: MHL', 
                                         'di: LLHH', 'tri: LMH')),
         sandhi_category = factor(sandhi_category,
                                  levels = c('Rise-fall', 'L-to-H Level',
                                             'LLH')),
         diortri = ifelse(diortri == 'di', 'disyllable', 'trisyllable'))

p_dist <-distri_prop2(f0_yas_comp1, diortri, sandhi_category)+
  theme(legend.title = element_blank())+
  scale_fill_manual(values = c( "#DDCC77", "#9B72C7", "#CC6677"))+
  labs(title = "Distribution of sandhi categories")+
  xlab("");p_dist
```
combination of plots
```{r warning=FALSE}
p <- ggarrange(
  ggarrange(p_di+ rremove("xlab"), 
            p_tri + rremove("ylab") + rremove("xlab"), 
            ncol = 2, widths = c(1, 1),
            labels = c('A', 'B'),
            font.label = list(size = 25, family = "Times")),
  ggarrange(p_comp + rremove("xlab"), 
            p_dist + rremove("ylab") + rremove("xlab"), 
            ncol = 2, widths = c(1.5, 1),
            labels = c('C', 'D'),
            font.label = list(size = 25, family = "Times")),
  nrow = 2, align = "v", heights = c(1.2, 1.2, 1.2)
  #common.legend = TRUE, legend = "top"
); p

ggsave(filename = file.path(plot_dir, "yash_di_comp.jpeg"),width = 10, height = 8)
```

## Distribution analysis

distributions
```{r}
f0_tri_ct_yas <- f0_tri_ct_yas %>% 
  mutate(sandhi_tone = factor(sandhi_tone, levels = c('LHL', 'LMH', 'LLH')))

# by syntax
p_syntax <- f0_tri_ct_yas %>% distri_prop2(syntax, sandhi_tone)+
  scale_fill_manual(values = c("#DDCC77","#9B72C7", "#CC6677"))+
  labs(title = "Distribution by syntax");p_syntax

# by syntax & tone 1
p_syntax_tone1 <- f0_tri_ct_yas %>% distri_prop(sync_tone1, sandhi_tone, syntax,
                             label_format = "paste0(round(prop * 100), '%')")+
  scale_fill_manual(values = c("#4477AA", "#117733", "#DDCC77", "#CC6677"))+
  theme(legend.title = element_blank())+
  labs(title = "Distribution by syntax and σ1 tone", 
       x = "initial tone")+
  geom_text(aes(x = sync_tone1, y = 1.05, label = count),
              size = 5, family = 'Times New Roman', inherit.aes = FALSE)+
  annotate("text", x = -Inf, y = 1.08, label = "n =",
             hjust = -0.1, vjust = 0, size = 5,  family = "Times New Roman");p_syntax_tone1

# by syntax & tones  
p_syntax_tone <- f0_tri_ct_yas %>% distri_prop(hist_tone23_mapped, sandhi_tone, syntax,
                             label_format = "paste0(round(prop * 100), '%')")+
  scale_fill_manual(values = c("#DDCC77","#9B72C7", "#CC6677"))+
  labs(title = "Distribution by syntax and σ2_σ3 tones", 
       x = "second tone_third tone")+
  scale_x_discrete(labels = setNames(f0_tri_ct_yas$tone23_colored,
                                     f0_tri_ct_yas$hist_tone23_mapped)) +
  theme(axis.text.x = element_markdown(angle = 45, hjust = 1))+
  geom_text(aes(x = hist_tone23_mapped, y = 1.05, label = count),
              size = 5, family = 'Times New Roman', inherit.aes = FALSE)+
  annotate("text", x = -Inf, y = 1.08, label = "n =",
             hjust = -0.1, vjust = 0, size = 5,  family = "Times New Roman");p_syntax_tone

p <- ggarrange(ggarrange(p_syntax, p_syntax_tone1,
                         widths = c(1, 2),
                         labels = c('A', 'B'),
                         font.label = list(size = 25, family = "Times")), 
            p_syntax_tone+ rremove("ylab"),
            nrow = 2,
            labels = c('', 'C'),
            font.label = list(size = 25, family = "Times"));p
 
 ggsave(filename = file.path(plot_dir, "yash_syntax.jpeg"),width = 10, height = 9)
```


```{r}
# by speaker
p_di_speaker <- f0_yas_comp1 %>% 
  filter(diortri == 'trisyllable') %>% 
  distri_prop2(speaker, sandhi_category, label_format = "paste0(round(prop * 100), '%')")+
  scale_fill_manual(values = c( "#DDCC77","#9B72C7", "#CC6677"))+
  labs(title = "Distribution by speaker in trisyllables");p_di_speaker

p_tri_speaker <- f0_yas_comp1 %>% 
  filter(diortri == 'disyllable') %>% 
  distri_prop2(speaker, sandhi_category, label_format = "paste0(round(prop * 100), '%')")+
  scale_fill_manual(values = c("#DDCC77","#9B72C7", "#CC6677"))+
  labs(title = "Distribution by speaker in disyllables");p_tri_speaker

p <- ggarrange(p_di_speaker, 
            p_tri_speaker , 
            widths = c(1, 1),
            labels = c('A', 'B'),
            font.label = list(size = 25, family = "Times"),
            common.legend = TRUE,
            legend = 'right');p

ggsave(filename = file.path(plot_dir, "yash_speaker.jpeg"),width = 10, height = 4.5)
```


```{r}
p_sub_cluster(f0_tri_ct_yas, syntax, sandhi_tone)
```


# yangping-initial

```{r}
p_cluster(f0_tri_ct_yap, sandhi_tone)+
  scale_color_manual(values = c("#CC6677"))
ggsave(filename = file.path(plot_dir, "yapi_clusters.jpeg"),width = 6, height = 4)
```

```{r message=FALSE, warning=FALSE}
# yangping-initial disyllabic dataset
f0_di_ct_yap <- f0_all_ct %>% 
  filter(syntax %in% c('L', 'M') & diortri == 'di' & hist_tone1 == 'yangping') %>% 
  mutate(sandhi_tone = case_when(sandhi_tone == 'HLLM' ~ 'HMML',
                                 sandhi_tone == 'LLRF' ~ 'LLHL', 
                                 sandhi_tone == 'LLLM' ~ 'LLLH',
                                 sandhi_tone == 'LMML' ~ 'LHML',
                                 .default = sandhi_tone)) %>% 
  filter(!ind_no %in% c('S2_1_ct', 'S2_11_ct', 'S2_27_ct', 'S3_5_ct', 'S3_19_ct', 'S5_27_ct',
                        'S2_44_ct', 'S3_37_ct', 'S3_44_ct', 'S6_16_ct', 'S6_31_ct', 'S6_39_ct', 'S7_33_ct')) %>% 
  filter(is.na(sandhi_tone) == FALSE) 

unique(f0_di_ct_yap$sandhi_tone) # check sandhi categories

# plotting
p_di <- p_cluster_cont(f0_di_ct_yap, sandhi_tone)+
  theme(legend.title = element_blank())+
  labs(title = 'Disyllabic sandhi');p_di
p_tri <- p_cluster_cont(f0_tri_ct_yap, sandhi_tone)+
  theme(legend.title = element_blank())+
  scale_color_manual(values = c("#CC6677"))+
  labs(title = 'Trisyllabic sandhi'); p_tri

unique(f0_tri_ct_yap$hist_tone2)

# check individual contours
# ggplotly(draw_by(f0_di_ct_lcmh_h, 'sandhi_tone'), tooltip = c('text', 'x'))
```


direct comparions between di & tri sandhi
```{r message=FALSE, warning=FALSE}
unique(f0_di_ct_yap$sandhi_tone)

f0_tri_yap_comp <- f0_tri_ct_yap %>% 
  select(-sync_tone23, -hist_tone23, -hist_tone23_mapped, -tone2_mapped, -tone23_colored, -all_tone) %>% 
  mutate(sandhi_category = case_when(sandhi_tone == 'LLH' ~ 'Rise',
                                     .default = sandhi_tone),
         sandhi_tone = paste0('tri: ', sandhi_tone))

# direct comparisons of similar sandhi contours
f0_di_yap_comp <- f0_di_ct_yap %>% 
  mutate(sandhi_category = case_when(sandhi_tone == 'LLLH' ~ 'Rise',
                                     .default = sandhi_tone),
         sandhi_tone = paste0('di: ', sandhi_tone))

f0_yap_comp <- rbind(f0_tri_yap_comp %>% filter(sandhi_tone %in% c('tri: LLH')),
                    f0_di_yap_comp %>% filter(sandhi_tone %in% c('di: LLLH'))) %>% 
  mutate(sandhi_tone = factor(sandhi_tone, 
                              levels = c('di: LLLH', 'tri: LLH')))
  
p_comp <- p_cluster_cont(f0_yap_comp, sandhi_tone, 'sandhi_category')+
   theme(legend.title = element_blank())+
  scale_color_manual(values = c("#CC6677", "#882255"))+
  labs(title = 'Comparisons of di- and tri-syllabic sandhi');p_comp
```


overall distribution comparison
```{r message=FALSE, warning=FALSE}
f0_yap_comp1 <- rbind(f0_tri_yap_comp,f0_di_yap_comp) %>% 
  mutate(sandhi_tone = factor(sandhi_tone, 
                              levels = c('di: LLLH', 'tri: LLH')),
         sandhi_category = factor(sandhi_category,
                                  levels = c('Rise', 'LLHH', 'LHML', 'LLHL')),
         diortri = ifelse(diortri == 'di', 'disyllable', 'trisyllable'))

p_dist <- distri_prop2(f0_yap_comp1, diortri, sandhi_category,
             label_format = "paste0(round(prop * 100), '% (', n, ')')")+
   theme(legend.title = element_blank())+
  scale_fill_manual(values = c( "#CC6677", "#117733", "#4477AA","#DDCC77", "#66CCEE"))+
  labs(title = "Distribution of sandhi categories")+
  xlab("");p_dist 
```


combination of plots
```{r warning=FALSE}
p <- ggarrange(
  ggarrange(p_di+ rremove("xlab"), 
            p_tri + rremove("ylab") + rremove("xlab"), 
            ncol = 2, widths = c(1, 1),
            labels = c('A', 'B'),
            font.label = list(size = 25, family = "Times")),
  ggarrange(p_comp + rremove("xlab"), 
            p_dist + rremove("ylab") + rremove("xlab"), 
            ncol = 2, widths = c(1.5, 1),
            labels = c('C', 'D'),
            font.label = list(size = 25, family = "Times")),
  nrow = 2, align = "v", heights = c(1.2, 1.2, 1.2)
  #common.legend = TRUE, legend = "top"
); p

ggsave(filename = file.path(plot_dir, "yapi_di_comp.jpeg"),width = 10, height = 8)
```