---
title: "sandhi_tri_yinshang"
author: "Yibing Shi"
date: "2024-12-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggtext)
library(grid)
library(plotly)
library(kml)
```

# Data preparations  
```{r data-construction, cache=FALSE}
load("XSTSF_production.RData")
source('functions.R')
```

paths
```{r}
current_dir <- getwd()
plot_dir <- file.path(current_dir, "plots")
```

datasets
```{r, cache=FALSE}
f0_all_ct <- f0_all_pre %>% filter(focus_condition == 'ct' ) %>% 
  group_by(speaker) %>% 
  mutate(norm_f0 = scale(log(f0))) %>% 
  ungroup() %>% 
  mutate(
    time = as.numeric(time),
    syllable_no = case_when(
    time > 0 & time < 11 ~ 1,
    time > 10 & time < 21 ~ 2,
    time > 20 & time < 31 ~ 3
  ),
  sync_tone1 = ifelse(sync_tone1 == 'RF', 'LHL', sync_tone1),
  sync_tone2 = ifelse(sync_tone2 == 'RF', 'LHL', sync_tone2),
  sync_tone3 = ifelse(sync_tone3 == 'RF', 'LHL', sync_tone3))

tone2_colors  <- c("YiPi" = "#E07B91", "YiSh" = "#8E735B", "YaPi" = "#6B8E66")
f0_tri_ct <- f0_all_ct
f0_tri_ct <- f0_all_ct %>% 
  filter(diortri == 'tri' & sandhi_tone != 'outlier')  %>% 
  mutate(across(c(hist_tone1, hist_tone2, hist_tone3), 
                ~ recode(.,
                         "yinping" = "YiPi",
                         "yinshang" = "YiSh",
                         "yinqu" = "YiQu",
                         "yangping" = "YaPi",
                         "yangshang" = "YaSh",
                         "yangqu" = "YaQu"))) %>% 
  mutate(sync_tone23 = paste0(sync_tone2, '_', sync_tone3),
         hist_tone23 = paste0(hist_tone2, '_', hist_tone3),
         hist_tone23_mapped = hist_tone23,
         all_tone = paste(sync_tone1, hist_tone23_mapped, sep = "_"),
         tone2_mapped = str_split_fixed(hist_tone23_mapped, "_", 2)[, 1],
         tone23_colored = paste0("<span style='color:", 
                                tone2_colors[f0_tri_ct$tone2_mapped], ";'>",
                                f0_tri_ct$tone2_mapped,  "</span>_<span>", 
                                str_split_fixed(f0_tri_ct$hist_tone23_mapped, "_", 2)[, 2],
                                "</span>"),
         sandhi_tone = ifelse(sandhi_tone == 'HHL', 'MML', 
                              ifelse(sandhi_tone == 'LHL', 'MHL', sandhi_tone)),
         syntax = case_when(syntax == '1+2MH' ~ '1+2AN',
                            syntax == '2+1VL' ~ '2+1VN',
                            syntax == '2+1LL' ~ '2+1NN',
                            .default = syntax)) %>% 
  mutate(syntax = factor(syntax, levels = c('1+2AN', '2+1NN', '1+2VO')))
  
f0_tri_ct_yp <- f0_tri_ct %>% filter(hist_tone1 == 'YiPi') 
f0_tri_ct_yap <- f0_tri_ct %>% filter(hist_tone1 == 'YaPi')
f0_tri_ct_ys <- f0_tri_ct %>% 
  filter(hist_tone1 == 'YiSh' & ! ind_no %in% c('S6_3_ct', 'S7_3_ct')) 
f0_tri_ct_yas <- f0_tri_ct %>% filter(hist_tone1 == 'YaSh')
```

## monosyllabic tones
```{r message=FALSE,warning=FALSE}
p1 <- f0_tri_ct_ys %>%
  distri_prop(hist_tone1, sync_tone1, syntax)+
  theme(axis.text.x = element_blank())+
  theme(plot.margin = margin(b = 20),
        legend.title = element_blank())+
  labs(fill = "synchronic tone",
       title = 'σ1 (yinshang)');p1
  
p2 <- f0_tri_ct_ys %>%
  filter(startsWith(hist_tone2, "Yi")) %>%
  distri_prop(hist_tone2, sync_tone2, syntax)+
  theme(plot.margin = margin(b = 20),
        legend.title = element_blank())+
  labs(fill = "synchronic tone",
       title = 'σ2 (High-register)');p2

p3 <- f0_tri_ct_ys %>%
  filter(startsWith(hist_tone2, "Ya")) %>%
  distri_prop(hist_tone2, sync_tone2, syntax)+
  scale_fill_manual(values = c('#C77CFF', '#7CAE00'))+
  theme(plot.margin = margin(b = 20),
        legend.title = element_blank())+
  labs(fill = "synchronic tone",
       title = 'σ2 (Low-register)');p3

p4 <- f0_tri_ct_ys %>%
  filter(startsWith(hist_tone3, "Yi")) %>%
  distri_prop(hist_tone3, sync_tone3, syntax)+
  theme(plot.margin = margin(b = 20),
        legend.title = element_blank())+
  labs(fill = "synchronic tone",
       title = 'σ3 (High-register)'); p4

p5 <- f0_tri_ct_ys %>%
  filter(startsWith(hist_tone3, "Ya")) %>%
  distri_prop(hist_tone3, sync_tone3, syntax)+
  theme(legend.title = element_blank())+
  scale_fill_manual(values = c('#C77CFF', '#7CAE00'))+
  theme(plot.margin = margin(b = 20))+
  labs(fill = "synchronic tone",
       title = 'σ3 (Low-register)');p5
```

combination
```{r warning=FALSE}
p <- ggarrange(
  ggarrange(p1 + rremove("ylab") + rremove("xlab"), NULL, 
            ncol = 2, widths = c(1, 1), 
            labels = c('A', ''),
            font.label = list(size = 25, family = "Times")),
  ggarrange(p2 + rremove("ylab") + rremove("xlab"), 
            p3 + rremove("ylab") + rremove("xlab"), 
            ncol = 2, widths = c(1, 1), 
            labels = c('B-1', 'B-2'),
            font.label = list(size = 25, family = "Times")),
  ggarrange(p4 + rremove("ylab") + rremove("xlab"), 
            p5 + rremove("ylab") + rremove("xlab"), 
            ncol = 2, widths = c(1, 1), 
            labels = c('C-1', 'C-2'),
            font.label = list(size = 25, family = "Times")),
  nrow = 3, align = "v", heights = c(1.2, 1.2, 1.2),
  common.legend = TRUE, legend = "top"
); p


p_annotated <- annotate_figure(
  p,
  bottom = text_grob("Historical tonal category", size = 20, family = "Times")
); p_annotated
     

ggsave(filename = file.path(plot_dir, "yish_mono.jpeg"),width = 10, height = 12)
```


# Sandhi categorisation

Auditory categorisation
```{r message=FALSE, warning=FALSE}
unique(f0_tri_ct_ys$sandhi_tone)

p_human <-p_cluster(f0_tri_ct_ys, sandhi_tone)+
  theme(legend.position = 'top',
        legend.title = element_blank());p_human
```

comparisons with yinping-initial sandhi
```{r}
unique(f0_tri_ct_yp$sandhi_tone)
f0_tri_yin_comp <- rbind(f0_tri_ct_yp %>% filter(sandhi_tone != 'MML'),
                         f0_tri_ct_ys %>% filter(sandhi_tone != 'MMM'))
p_cluster_cont(f0_tri_yin_comp, hist_tone1, 'sandhi_tone')
```


*k*-means clustering
```{r}
# data preparation
f0_tri_ct_ys_kmeans <- f0_tri_ct_ys %>% 
  select(-diortri, -syllable_no, -focus_no, -f0) %>% 
  spread(time, norm_f0)

# k-means clustering
cluster_model <- k_means_clustering(f0_tri_ct_ys_kmeans)
kml(cluster_model, nbClusters = 2:10) 
plotAllCriterion(cluster_model)
kml::plot(cluster_model, 3, parTraj=parTRAJ(col="clusters"))
kml::plot(cluster_model, 5, parTraj=parTRAJ(col="clusters"))


# get cluster results
f0_tri_ct_ys_kmeans <- f0_tri_ct_ys_kmeans %>% 
  mutate(cluster5 = getClusters(cluster_model, 5))

f0_tri_ct_ys_kmeans <- wide_to_long(f0_tri_ct_ys_kmeans) %>% 
  mutate(syllable_no = ifelse(time < 11, '1', 
                              ifelse(time < 21, '2', '3'))) 
```


*k*-means cluster visualisation
```{r warning=FALSE, message=FALSE}
f0_tri_ct_ys_kmeans <- f0_tri_ct_ys_kmeans %>% 
  mutate(cluster5_reorder = case_when(cluster5 == 'D' ~ 'A',
                                      cluster5 == 'C' ~ 'B',
                                      cluster5 == 'A' ~ 'C',
                                      cluster5 == 'E' ~ 'D',
                                      cluster5 == 'B' ~ 'E'))

p_kmeans5 <- p_cluster(f0_tri_ct_ys_kmeans, cluster5_reorder)+
  theme(legend.position = 'top',
        legend.title = element_blank());p_kmeans5
```

heatmap distribution
```{r message=FALSE, warning=FALSE}
heatmap_df <- heatmap_data(f0_tri_ct_ys_kmeans, cluster5_reorder) 
p_htmap5 <- compare_cluster(heatmap_df, 'cluster5_reorder'); p_htmap5
```
combination

```{r warning=FALSE}
p <- ggarrange(
  p_human+ labs(title = 'Perceptual clusters'), 
  p_kmeans5+ rremove("ylab") +labs(title = 'k-means clusters \n(k=5)'), 
  p_htmap5 + rremove("ylab") + rremove("xlab")+labs(title = 'Perceptual vs. K-means \n(k=5)'),
  nrow = 1, heights = c(1, 1, 1),
  labels = c('A', 'B', 'C'),
  font.label = list(size = 25, family = "Times",hjust = -1),
  align = "hv",
  legend = "bottom" 
);p


ggsave(filename = file.path(plot_dir, "yish_clusters.jpeg"),width = 12, height = 4)
```


# Comparisons with disyllabic sandhi

plotting sandhi patterns
```{r message=FALSE, warning=FALSE}
# yinshang-initial disyllabic dataset
f0_di_ct_ys <- f0_all_ct %>% 
  filter(syntax %in% c('L', 'M') & diortri == 'di' & hist_tone1 == 'yinshang') %>% 
  mutate(sandhi_tone = case_when(sandhi_tone == 'HMML' ~ 'HHML',
                                 .default = sandhi_tone)) %>% 
  filter(!ind_no %in% c('S2_1_ct', 'S2_11_ct', 'S2_27_ct', 
                        'S3_5_ct', 'S3_19_ct', 'S5_27_ct')) %>% 
  filter(is.na(sandhi_tone) == FALSE,
         sandhi_tone != 'MMMH') 

unique(f0_di_ct_ys$sandhi_tone) # check sandhi categories

# plotting
p_di <- p_cluster_cont(f0_di_ct_ys, sandhi_tone)+
  theme(legend.title = element_blank())+
  scale_color_manual(values = c("#9B72C7", "#DDCC77"))+
  labs(title = 'Disyllabic sandhi'); p_di
p_tri <- p_cluster_cont(f0_tri_ct_ys, sandhi_tone)+
  theme(legend.title = element_blank())+
  labs(title = 'Trisyllabic sandhi'); p_tri

# check individual contours
# ggplotly(draw_by(f0_di_ct_lcmh_h, 'sandhi_tone'), tooltip = c('text', 'x'))
```


direct comparions between di & tri sandhi
```{r message=FALSE, warning=FALSE}
f0_tri_ys_comp <- f0_tri_ct_ys %>% 
  select(-sync_tone23, -hist_tone23, -hist_tone23_mapped, -tone2_mapped, -tone23_colored, -all_tone) %>% 
  mutate(sandhi_category = case_when(sandhi_tone == 'MHL' ~ 'Rise-fall',
                                     sandhi_tone == 'MMM' ~ 'Level',
                                     .default = sandhi_tone),
         sandhi_tone = paste0('tri: ', sandhi_tone))

# direct comparisons of similar sandhi contours
f0_di_ys_comp <- f0_di_ct_ys %>% 
  mutate(sandhi_category = case_when(sandhi_tone == 'HHHH' ~ 'Level',
                                     sandhi_tone == 'MHHL' ~ 'Rise-fall',
                                     .default = 'Rise'),
         sandhi_tone = paste0('di: ', sandhi_tone))

unique(f0_tri_ys_comp$sandhi_tone)

f0_ys_comp <- rbind(f0_tri_ys_comp %>% 
                      filter(sandhi_tone %in% c('tri: MMM', 'tri: MHL')),
                    f0_di_ys_comp) %>% 
  mutate(sandhi_tone = factor(sandhi_tone, 
                              levels = c('di: HHHH', 'tri: MMM',
                                         'di: MHHL', 'tri: MHL')))
  
p_comp <- p_cluster_cont(f0_ys_comp, sandhi_tone, 'sandhi_category')+
  theme(legend.title = element_blank(),
        legend.position = 'bottom')+
  scale_color_manual(values = c("#CC99CC", "#9B72C7", "#DDCC77", "#AA9944"))+
  labs(title = 'Comparisons of di- and tri-syllabic sandhi');p_comp
```

overall distribution comparison
```{r message=FALSE, warning=FALSE}
f0_ys_comp1 <- rbind(f0_tri_ys_comp,f0_di_ys_comp) %>% 
  mutate(sandhi_tone = factor(sandhi_tone, 
                              levels = c('di: MHHL', 'tri: MHL',
                                         'di: HHHH', 'tri: MMM')),
         sandhi_category = factor(sandhi_category,
                                  levels = c('Rise-fall', 'Level',
                                             'HLM', 'HML', 'MMH')),
         diortri = ifelse(diortri == 'di', 'disyllable', 'trisyllable'))

p_dist <- distri_prop2(f0_ys_comp1, diortri, sandhi_category,
                       label_format = "paste0(round(prop * 100), '% (', n, ')')")+
  theme(legend.title = element_blank())+
  scale_fill_manual(values = c("#DDCC77","#9B72C7","#4477AA", "#117733", "#CC6677" ))+
  labs(title = "Frequency distribution")+
  xlab("");p_dist
```
combination of plots
```{r warning=FALSE}
p <- ggarrange(
  ggarrange(p_di+ rremove("xlab"), 
            p_tri + rremove("ylab") + rremove("xlab"), 
            ncol = 2, widths = c(1, 1),
            labels = c('A', 'B'),
            font.label = list(size = 25, family = "Times")),
  ggarrange(p_comp + rremove("xlab"), 
            p_dist + rremove("ylab") + rremove("xlab"), 
            ncol = 2, widths = c(1.5, 1),
            labels = c('C', 'D'),
            font.label = list(size = 25, family = "Times")),
  nrow = 2, align = "v", heights = c(1.2, 1.2, 1.2)
  #common.legend = TRUE, legend = "top"
); p

ggsave(filename = file.path(plot_dir, "yish_di_comp.jpeg"),width = 10, height = 8)
```


## Distribution analysis

distributions
```{r}
f0_tri_ct_ys <- f0_tri_ct_ys %>% 
  mutate(sandhi_tone = factor(sandhi_tone, levels = c('MHL', 'MMM', 'HLM', 'HML', 'MMH'))) %>% 
  mutate(syntax = factor(syntax, levels = c('1+2AN', '2+1NN', '2+1VN', '1+2VO')))
# by syntax
p_syntax <- f0_tri_ct_ys %>% distri_prop2(syntax, sandhi_tone)+
  scale_fill_manual(values = c("#DDCC77","#9B72C7","#4477AA", "#117733", "#CC6677"))+
  labs(title = "Distribution by syntax");p_syntax 

# by syntax & tones  
p_syntax_tone <- f0_tri_ct_ys %>% distri_prop(hist_tone23_mapped, sandhi_tone, syntax,
                             label_format = "paste0(round(prop * 100), '%')")+
  scale_fill_manual(values = c("#DDCC77","#9B72C7","#4477AA", "#117733", "#CC6677"))+
  labs(title = "Distribution by syntax and tones", 
       x = "second tone_third tone")+
  scale_x_discrete(labels = setNames(f0_tri_ct_ys$tone23_colored,
                                     f0_tri_ct_ys$hist_tone23_mapped)) +
  theme(axis.text.x = element_markdown(angle = 45, hjust = 1))+
  geom_text(aes(x = hist_tone23_mapped, y = 1.05, label = count),
              size = 5, family = 'Times New Roman', inherit.aes = FALSE)+
  annotate("text", x = -Inf, y = 1.08, label = "n =",
             hjust = -0.1, vjust = 0, size = 5,  family = "Times New Roman");p_syntax_tone

p <- ggarrange(p_syntax, 
            p_syntax_tone+ rremove("ylab") , 
            widths = c(1, 2.2),
            labels = c('A', 'B'),
            font.label = list(size = 25, family = "Times"),
            common.legend = TRUE,
            legend = 'right');p
 
 ggsave(filename = file.path(plot_dir, "yish_syntax.jpeg"),width = 10, height = 4.5)
```


```{r}
# by speaker
p_di_speaker <- f0_ys_comp1 %>% 
  filter(diortri == 'trisyllable') %>% 
  distri_prop2(speaker, sandhi_category, label_format = "paste0(round(prop * 100), '%')")+
  scale_fill_manual(values = c("#DDCC77","#9B72C7","#4477AA", "#117733", "#CC6677"))+
  labs(title = "Distribution by speaker in trisyllables");p_di_speaker

p_tri_speaker <- f0_ys_comp1 %>% 
  filter(diortri == 'disyllable') %>% 
  distri_prop2(speaker, sandhi_category, label_format = "paste0(round(prop * 100), '%')")+
  scale_fill_manual(values = c("#DDCC77","#9B72C7","#4477AA", "#117733", "#CC6677"))+
  labs(title = "Distribution by speaker in disyllables");p_tri_speaker

p <- ggarrange(p_di_speaker, 
            p_tri_speaker , 
            widths = c(1, 1),
            labels = c('A', 'B'),
            font.label = list(size = 25, family = "Times"),
            common.legend = TRUE,
            legend = 'right');p

ggsave(filename = file.path(plot_dir, "yish_speaker.jpeg"),width = 10, height = 4.5)
```



