---
title: "Trisyllabic sandhi in the Xiangshan dialect"
author: "Yibing Shi"
date: "2024-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(ggtext)
library(grid)
library(plotly)
library(kml)
```

# Data preparations  
```{r data-construction, cache=FALSE}
load("XSTSF_production.RData")
source('functions.R')
```

paths
```{r}
current_dir <- getwd()
plot_dir <- file.path(current_dir, "plots")
```


datasets
```{r, cache=FALSE}
f0_all_ct <- f0_all_pre %>% filter(focus_condition == 'ct' ) %>% 
  group_by(speaker) %>% 
  mutate(norm_f0 = scale(log(f0))) %>% 
  ungroup() %>% 
  mutate(
    time = as.numeric(time),
    syllable_no = case_when(
    time > 0 & time < 11 ~ 1,
    time > 10 & time < 21 ~ 2,
    time > 20 & time < 31 ~ 3
  ),
  sync_tone1 = ifelse(sync_tone1 == 'RF', 'LHL', sync_tone1),
  sync_tone2 = ifelse(sync_tone2 == 'RF', 'LHL', sync_tone2),
  sync_tone3 = ifelse(sync_tone3 == 'RF', 'LHL', sync_tone3))

tone2_colors  <- c("YiPi" = "#E07B91", "YiSh" = "#8E735B", "YaPi" = "#6B8E66")

f0_tri_ct <- f0_all_ct %>% filter(diortri == 'tri' & sandhi_tone != 'outlier')
f0_tri_ct <- f0_all_ct %>% 
  filter(diortri == 'tri' & sandhi_tone != 'outlier')  %>% 
  mutate(across(c(hist_tone1, hist_tone2, hist_tone3), 
                ~ recode(.,
                         "yinping" = "YiPi",
                         "yinshang" = "YiSh",
                         "yinqu" = "YiQu",
                         "yangping" = "YaPi",
                         "yangshang" = "YaSh",
                         "yangqu" = "YaQu"))) %>% 
  mutate(sync_tone23 = paste0(sync_tone2, '_', sync_tone3),
         hist_tone23 = paste0(hist_tone2, '_', hist_tone3),
         hist_tone23_mapped = hist_tone23,
          #hist_tone23_mapped = gsub("yinping", "YiPi", 
          #               gsub("yangping", "YaPi", 
          #               gsub("yinshang", "YiSh", 
          #               gsub("yangshang", "YaSh", 
          #               gsub("yinqu", "YiQu", 
          #               gsub("yangqu", "YaQu", hist_tone23)))))),
         all_tone = paste(sync_tone1, hist_tone23_mapped, sep = "_"),
         tone2_mapped = str_split_fixed(hist_tone23_mapped, "_", 2)[, 1],
         tone23_colored = paste0("<span style='color:", 
                          map_chr(tone2_mapped, ~ tone2_colors[.]), ";'>",
                          tone2_mapped, "</span>_<span>", 
                          str_split_fixed(hist_tone23_mapped, "_", 2)[, 2],
                          "</span>"),
         sandhi_tone = case_when(sandhi_tone == 'HHL' ~ 'MHL', 
                                 sandhi_tone == 'LHL' & grepl("^Yi", hist_tone1) ~ 'MHL',
                                 sandhi_tone %in% c('HML', 'MML') & syntax == '2+1VL' ~ 'MHL',
                                 .default = sandhi_tone),
         syntax = case_when(syntax == '1+2MH' ~ '1+2AN',
                            syntax == '2+1VL' ~ '2+1VN',
                            syntax == '2+1LL' ~ '2+1NN',
                            .default = syntax)) %>% 
  mutate(syntax = factor(syntax, levels = c('1+2AN', '2+1NN', '2+1VN', '1+2VO')))
  
f0_tri_ct_yp <- f0_tri_ct %>% filter(hist_tone1 == 'YiPi') 
f0_tri_ct_yap <- f0_tri_ct %>% filter(hist_tone1 == 'YaPi')
f0_tri_ct_ys <- f0_tri_ct %>% filter(hist_tone1 == 'YiSh')
f0_tri_ct_yas <- f0_tri_ct %>% filter(hist_tone1 == 'YaSh')
```


# Monosyllabic tones

```{r warning=FALSE}
p1 <- f0_tri_ct_yp %>%
  distri_prop(hist_tone1, sync_tone1, syntax)+
  theme(axis.text.x = element_blank(),
        legend.title = element_blank())+
  theme(plot.margin = margin(b = 20))+
  labs(fill = "synchronic tone",
       title = 'σ1 (yinping)'); p1
  
p2 <- f0_tri_ct_yp %>%
  filter(startsWith(hist_tone2, "Yi")) %>%
  distri_prop(hist_tone2, sync_tone2, syntax)+
  theme(plot.margin = margin(b = 20),
        legend.title = element_blank())+
  labs(fill = "synchronic tone",
       title = 'σ2 (High-register)'); p2

p3 <- f0_tri_ct_yp %>%
  filter(startsWith(hist_tone2, "Ya")) %>%
  distri_prop(hist_tone2, sync_tone2, syntax)+
  theme(plot.margin = margin(b = 20),
        legend.title = element_blank())+
  labs(fill = "synchronic tone",
       title = 'σ2 (Low-register)')+
  scale_fill_manual(values = c('#7CAE00', '#C77CFF')); p3

p4 <- f0_tri_ct_yp %>%
  filter(startsWith(hist_tone3, "Yi")) %>%
  distri_prop(hist_tone3, sync_tone3, syntax)+
  theme(legend.title = element_blank())+
  labs(fill = "synchronic tone",
       title = 'σ3 (High-register)'); p4

p5 <- f0_tri_ct_yp %>%
  filter(startsWith(hist_tone3, "Ya")) %>%
  distri_prop(hist_tone3, sync_tone3, syntax)+
  theme(legend.title = element_blank())+
  labs(fill = "synchronic tone",
       title = 'σ2 (Low-register)')+
  scale_fill_manual(values = c('#7CAE00', '#C77CFF')); p5
```


```{r warning=FALSE}
p <- ggarrange(
  ggarrange(p1 + rremove("ylab") + rremove("xlab"), NULL, 
            ncol = 2, widths = c(1, 1), 
            labels = c('A', ''),
            font.label = list(size = 25, family = "Times")),
  ggarrange(p2 + rremove("ylab") + rremove("xlab"), 
            p3 + rremove("ylab") + rremove("xlab"), 
            ncol = 2, widths = c(1, 1), 
            labels = c('B-1', 'B-2'),
            font.label = list(size = 25, family = "Times")),
  ggarrange(p4 + rremove("ylab") + rremove("xlab"), 
            p5 + rremove("ylab") + rremove("xlab"), 
            ncol = 2, widths = c(1, 1), 
            labels = c('C-1', 'C-2'),
            font.label = list(size = 25, family = "Times")),
  nrow = 3, align = "v", heights = c(1.2, 1.2, 1.2),
  common.legend = TRUE, legend = "top"
); p


p_annotated <- annotate_figure(
  p,
  bottom = text_grob("Historical tonal category", size = 20, family = "Times")
); p_annotated
     

ggsave(filename = file.path(plot_dir, "yipi_mono.jpeg"),width = 10, height = 12)
```



# Sandhi categorisation

Auditory categorisation
```{r}
unique(f0_tri_ct_yp$sandhi_tone)

p_human <- p_cluster(f0_tri_ct_yp, sandhi_tone);p_human
```

*k*-means clustering
```{r}
# data preparation
f0_tri_ct_yp_kmeans <- f0_tri_ct_yp %>% 
  select(-diortri, -syllable_no, -focus_no, -f0) %>% 
  spread(time, norm_f0)

# k-means clustering
cluster_model <- k_means_clustering(f0_tri_ct_yp_kmeans)
kml(cluster_model, nbClusters = 2:10) 
kml::plot(cluster_model, 2, parTraj=parTRAJ(col="clusters"))
kml::plot(cluster_model, 3, parTraj=parTRAJ(col="clusters"))
kml::plot(cluster_model, 4, parTraj=parTRAJ(col="clusters"))
kml::plot(cluster_model, 5, parTraj=parTRAJ(col="clusters"))
plotAllCriterion(cluster_model)


# get cluster results
f0_tri_ct_yp_kmeans <- f0_tri_ct_yp_kmeans %>% 
  mutate(cluster3 = getClusters(cluster_model, 3),
         cluster4 = getClusters(cluster_model, 4),
         cluster5 = getClusters(cluster_model, 5))

f0_tri_ct_yp_kmeans <- wide_to_long(f0_tri_ct_yp_kmeans) %>% 
  mutate(syllable_no = ifelse(time < 11, '1', 
                              ifelse(time < 21, '2', '3'))) 
```

*k*-means cluster visualisation
```{r warning=FALSE, message=FALSE}
f0_tri_ct_yp_kmeans <- f0_tri_ct_yp_kmeans %>% 
  mutate(cluster3_reorder = case_when(cluster3 == 'A' ~ 'A',
                                      cluster3 == 'C' ~ 'B',
                                      cluster3 == 'B' ~ 'C'),
         cluster4_reorder = case_when(cluster4 == 'B' ~ 'A',
                                      cluster4 == 'A' ~ 'B',
                                      cluster4 == 'D' ~ 'C',
                                      cluster4 == 'C' ~ 'D'))

p_kmeans3 <- p_cluster(f0_tri_ct_yp_kmeans, cluster3_reorder)+
  scale_color_manual(values = c("#4477AA", "#DDCC77", "#CC6677"));p_kmeans3
p_kmeans4 <- p_cluster(f0_tri_ct_yp_kmeans, cluster4_reorder);p_kmeans4
```

heatmap distribution
```{r message=FALSE, warning=FALSE}
heatmap_df <- heatmap_data(f0_tri_ct_yp_kmeans, cluster3_reorder) 
p_htmap3 <- compare_cluster(heatmap_df, 'cluster3_reorder'); p_htmap3

heatmap_df <- heatmap_data(f0_tri_ct_yp_kmeans, cluster4_reorder) 
p_htmap4 <- compare_cluster(heatmap_df, 'cluster4_reorder'); p_htmap4
```

combinations of plots

```{r warning=FALSE}
p_col1 <- ggarrange(
  p_human+ rremove("ylab") + rremove("xlab")+labs(title = 'Perceptual clusters'), 
  p_kmeans4+ rremove("ylab") + rremove("xlab")+labs(title = 'k-means clusters (k=4)'), 
  p_kmeans3+ rremove("ylab") + rremove("xlab")+labs(title = 'k-means clusters (k=3)'), 
  ncol = 1, heights = c(1.2, 1.2, 1.2),
  labels = c('A', 'B-1', 'C-1'),
  font.label = list(size = 25, family = "Times",hjust = -1),
  align = "v"
)

p_col1 <- annotate_figure(
  p_col1,
  left = text_grob("Normalised f0", rot = 90, size = 20, family = "Times"),
  bottom = text_grob("Normalised time", size = 20, family = "Times")
)

p_col2 <- ggarrange(
  NULL, 
  p_htmap4 + rremove("ylab") + rremove("xlab")+labs(title = 'Perceptual vs. K-means \n(k=4)'), 
  p_htmap3 + rremove("ylab") + rremove("xlab")+labs(title = 'Perceptual vs. K-means \n(k=3)'), 
  ncol = 1, heights = c(1.2, 1.2, 1.2),
  labels = c('', 'B-2', 'C-2'),
  font.label = list(size = 25, family = "Times", hjust = -1),
  common.legend = TRUE,
  legend = 'bottom'
)

p_clusters <- ggarrange(
  p_col1, p_col2, 
  ncol = 2, widths = c(1.4, 1), align = "v"
)

p_clusters
ggsave(filename = file.path(plot_dir, "yipi_clusters.jpeg"),width = 10, height = 11)
```




# Comparisons with disyllabic sandhi

H-M-L system validation
```{r}
# monosyllable
unique(f0_all_ct$diortri) # check diortri value for monosyllables
f0_mono_ct <- f0_all_ct %>% 
  filter(is.na(diortri) == TRUE) %>% 
  mutate(syllable_no = 1) # select monosyllabic data
unique(f0_mono_ct$token) # check if dataset is correct
unique(f0_mono_ct$citation_tone) # check tone inventory 
unique(f0_mono_ct$syllable_no) 

p_cluster(f0_mono_ct, citation_tone)
p_cluster(f0_mono_ct, citation_tone, 'speaker')

# S6's mono tones
f0_mono_s6 <- filter(f0_mono_ct, speaker == 'S6')
p_cluster(f0_mono_s6, citation_tone)
```

plotting sandhi patterns
```{r message=FALSE, warning=FALSE}
# yinping-initial disyllabic dataset
f0_di_ct_yp <- f0_all_ct %>% 
  filter(syntax %in% c('L', 'M') & diortri == 'di' & hist_tone1 == 'yinping') %>% 
  mutate(sandhi_tone = case_when(sandhi_tone == 'HLLM' ~ 'HHML',
                                 sandhi_tone == 'HMML' ~ 'HHML',
                                 sandhi_tone == 'LLHL' ~ 'LLRF', 
                                 .default = sandhi_tone)) %>% 
  filter(!ind_no %in% c('S2_1_ct', 'S2_11_ct', 'S2_27_ct', 
                        'S3_5_ct', 'S3_19_ct', 'S5_27_ct')) %>% 
  filter(is.na(sandhi_tone) == FALSE,
         sandhi_tone != 'HHHH') 

unique(f0_di_ct_yp$sandhi_tone) # check sandhi categories

# plotting
p_di <- p_cluster_cont(f0_di_ct_yp, sandhi_tone)+
  theme(legend.title = element_blank())+
  scale_color_manual(values = c("#4477AA", "#DDCC77", "#CC6677"))+
  labs(title = 'Disyllabic sandhi'); p_di
p_tri <- p_cluster_cont(f0_tri_ct_yp, sandhi_tone)+
  theme(legend.title = element_blank())+
  labs(title = 'Trisyllabic sandhi'); p_tri

# check individual contours
# ggplotly(draw_by(f0_di_ct_lcmh_h, 'sandhi_tone'), tooltip = c('text', 'x'))
```

direct comparions between di & tri sandhi
```{r message=FALSE, warning=FALSE}
f0_tri_yp_comp <- f0_tri_ct_yp %>% 
  select(-sync_tone23, -hist_tone23, -hist_tone23_mapped, -tone2_mapped, -tone23_colored, -all_tone) %>% 
  mutate(sandhi_category = case_when(sandhi_tone == 'HLM' ~ 'Fall',
                                     sandhi_tone == 'MHL' ~ 'Rise-fall',
                                     sandhi_tone == 'MMH' ~ 'Rise',
                                     .default = sandhi_tone),
         sandhi_tone = paste0('tri: ', sandhi_tone))

# direct comparisons of similar sandhi contours
f0_di_yp_comp <- f0_di_ct_yp %>% 
  mutate(sandhi_category = case_when(sandhi_tone == 'HHML' ~ 'Fall',
                                     sandhi_tone == 'MHHL' ~ 'Rise-fall',
                                     .default = 'Rise'),
         sandhi_tone = paste0('di: ', sandhi_tone))

f0_yp_comp <- rbind(f0_tri_yp_comp %>% 
                      filter(sandhi_tone %in% c('tri: HLM', 'tri: MHL', 'tri: MMH')),
                    f0_di_yp_comp) %>% 
  mutate(sandhi_tone = factor(sandhi_tone, 
                              levels = c('di: HHML', 'tri: HLM', 
                                         'di: MMMH', 'tri: MMH',
                                         'di: MHHL', 'tri: MHL')))
  
p_comp <- p_cluster_cont(f0_yp_comp, sandhi_tone, 'sandhi_category')+
  theme(legend.position = 'bottom',
        legend.title = element_blank())+
  scale_color_manual(values = c("#4477AA", "#88D3E9","#CC6677", "#882255", "#DDCC77", "#AA9944"))+
  labs(title = 'Comparisons of di- and tri-syllabic sandhi'); p_comp
```

overall distribution comparison
```{r message=FALSE, warning=FALSE}
f0_yp_comp1 <- rbind(f0_tri_yp_comp,f0_di_yp_comp) %>% 
  mutate(sandhi_tone = factor(sandhi_tone, 
                              levels = c('di: HHML', 'tri: HLM', 
                                         'di: MMMH', 'tri: MMH',
                                         'di: MHHL', 'tri: MHL')),
         sandhi_category = factor(sandhi_category,
                                  levels = c('Fall', 'Rise', 'Rise-fall',
                                             'HML', 'MML')),
         diortri = ifelse(diortri == 'di', 'disyllable', 'trisyllable'))

p_dist <- distri_prop2(f0_yp_comp1, diortri, sandhi_category)+
  theme(legend.title = element_blank())+
  scale_fill_manual(values = c("#4477AA", "#CC6677", "#DDCC77", "#117733"))+
  labs(title = "Frequency distribution")+
  xlab(""); p_dist
```

combination of plots
```{r warning=FALSE}
p <- ggarrange(
  ggarrange(p_di+ rremove("xlab"), 
            p_tri + rremove("ylab") + rremove("xlab"), 
            ncol = 2, widths = c(1, 1),
            labels = c('A', 'B'),
            font.label = list(size = 25, family = "Times")),
  ggarrange(p_comp + rremove("xlab"), 
            p_dist + rremove("ylab") + rremove("xlab"), 
            ncol = 2, widths = c(1.5, 1),
            labels = c('C', 'D'),
            font.label = list(size = 25, family = "Times")),
  nrow = 2, align = "v", heights = c(1.2, 1.2, 1.2)
  #common.legend = TRUE, legend = "top"
); p

ggsave(filename = file.path(plot_dir, "yipi_di_comp.jpeg"),width = 10, height = 8)
```



## Distribution analysis

distributions
```{r warning=FALSE}
# by syntax
f0_tri_ct_yp <- f0_tri_ct_yp %>% 
  mutate(factor(sandhi_tone, levels = c('HLM', 'HML', 'MHL', 'MMH')))
p_syntax <- f0_tri_ct_yp %>%   
  distri_prop2(syntax, sandhi_tone)+
  scale_fill_manual(values = c("#4477AA",  "#117733", "#DDCC77", "#CC6677"))+
  theme(legend.position = 'top')+
  labs(title = "Distribution by syntax");p_syntax

# by syntax & initial tones
p_syntax_tone1 <- f0_tri_ct_yp %>% distri_prop(sync_tone1, sandhi_tone, syntax,
                             label_format = "paste0(round(prop * 100), '%')")+
  scale_fill_manual(values = c("#4477AA", "#117733", "#DDCC77", "#CC6677"))+
  theme(legend.title = element_blank())+
  labs(title = "Distribution by syntax and σ1 tone", 
       x = "initial tone")+
  geom_text(aes(x = sync_tone1, y = 1.05, label = count),
              size = 5, family = 'Times New Roman', inherit.aes = FALSE)+
  annotate("text", x = -Inf, y = 1.08, label = "n =",
             hjust = -0.1, vjust = 0, size = 5,  family = "Times New Roman");p_syntax_tone1

# by syntax & second_third tones  
p_syntax_tone <- f0_tri_ct_yp %>% distri_prop(hist_tone23, sandhi_tone, syntax,
                             label_format = "paste0(round(prop * 100), '%')")+
  scale_fill_manual(values = c("#4477AA", "#117733", "#DDCC77", "#CC6677"))+
  theme(legend.title = element_blank())+
  labs(title = "Distribution by syntax and σ2_σ3 tones", 
       x = "second tone_third tone")+
  scale_x_discrete(labels = setNames(f0_tri_ct_yp$tone23_colored,
                                     f0_tri_ct_yp$hist_tone23_mapped)) +
  theme(axis.text.x = element_markdown(angle = 45, hjust = 1))+
  geom_text(aes(x = hist_tone23, y = 1.05, label = count),
              size = 5, family = 'Times New Roman', inherit.aes = FALSE)+
  annotate("text", x = -Inf, y = 1.08, label = "n =",
             hjust = -0.1, vjust = 0, size = 5,  family = "Times New Roman");p_syntax_tone

p <- ggarrange(ggarrange(p_syntax, p_syntax_tone1,
                         widths = c(1, 2),
                         labels = c('A', 'B'),
                         font.label = list(size = 25, family = "Times")), 
            p_syntax_tone+ rremove("ylab"),
            nrow = 2,
            labels = c('', 'C'),
            font.label = list(size = 25, family = "Times"));p
 
 ggsave(filename = file.path(plot_dir, "yipi_syntax.jpeg"),width = 10, height = 9)
```


```{r warning=FALSE}
# by speaker
p_di_speaker <- f0_yp_comp1 %>% 
  filter(diortri == 'trisyllable') %>% 
  distri_prop2(speaker, sandhi_category)+
  scale_fill_manual(values = c("#4477AA", "#CC6677", "#DDCC77", "#117733"))+
  labs(title = "Distribution by speaker in trisyllables");p_di_speaker

p_tri_speaker <-f0_yp_comp1 %>% 
  filter(diortri == 'disyllable') %>% 
  distri_prop2(speaker, sandhi_category)+
  scale_fill_manual(values = c("#4477AA", "#CC6677", "#DDCC77"))+
  labs(title = "Distribution by speaker in disyllables");p_tri_speaker

 p <- ggarrange(p_di_speaker, 
            p_tri_speaker , 
            widths = c(1, 1),
            labels = c('A', 'B'),
            font.label = list(size = 25, family = "Times"),
            common.legend = TRUE,
            legend = 'right');p
 
 ggsave(filename = file.path(plot_dir, "yipi_speaker.jpeg"),width = 10, height = 4.5)
```


```{r}
p_sub_cluster(f0_tri_ct_yp, syntax, sandhi_tone)
```

try building model
```{r eval=FALSE}
f0_tri_ct_yp <- f0_tri_ct_yp %>% 
  mutate(sandhi_tone = as.factor(sandhi_tone),
         sync_tone1 = as.factor(sync_tone1),
         hist_tone2 = as.factor(hist_tone2),
         hist_tone3 = as.factor(hist_tone3),
         syntax = as.factor(syntax),
         speaker = as.factor(speaker))

model <- multinom(sandhi_tone ~ sync_tone1 + hist_tone2 + hist_tone3 + syntax + speaker, data = f0_tri_ct_yp)
summary(model)

model_a1 <- multinom(sandhi_tone ~ sync_tone1 + hist_tone2 + hist_tone3 + speaker*syntax, data = f0_tri_ct_yp)
anova(model_a1, model, test = "Chisq")

model_1 <- multinom(sandhi_tone ~ hist_tone2 + hist_tone3 + syntax + speaker, data = f0_tri_ct_yp)
anova(model_1, model, test = "Chisq")

model_2 <- multinom(sandhi_tone ~ sync_tone1 + hist_tone3 + syntax + speaker, data = f0_tri_ct_yp)
anova(model_2, model, test = "Chisq")

model_3 <- multinom(sandhi_tone ~ sync_tone1 + hist_tone2 + syntax + speaker, data = f0_tri_ct_yp)
anova(model_3, model, test = "Chisq")

model_4 <- multinom(sandhi_tone ~ sync_tone1 + hist_tone2 + hist_tone3 + syntax, data = f0_tri_ct_yp)
anova(model_4, model, test = "Chisq")

model_5 <- multinom(sandhi_tone ~ sync_tone1 + hist_tone2 + hist_tone3 + speaker, data = f0_tri_ct_yp)
anova(model_5, model, test = "Chisq")
```



## yangshang-initial
```{r}
unique(f0_tri_ct_yas$sandhi_tone)

p_cluster(f0_tri_ct_yas, sandhi_tone)
```

monosyllabic tones
```{r}
f0_tri_ct_yas %>%
  distri_prop(hist_tone1, sync_tone1, syntax)+
  labs(fill = "synchronic tone")
  
f0_tri_ct_yas %>%
  filter(startsWith(hist_tone2, "yin")) %>%
  distri_prop(hist_tone2, sync_tone2, syntax)+
  labs(fill = "synchronic tone")

f0_tri_ct_yas %>%
  filter(startsWith(hist_tone3, "yin")) %>%
  distri_prop(hist_tone3, sync_tone3, syntax)+
  labs(fill = "synchronic tone")

f0_tri_ct_yas %>%
  filter(startsWith(hist_tone2, "yang")) %>%
  distri_prop(hist_tone2, sync_tone2, syntax)+
  labs(fill = "synchronic tone")

f0_tri_ct_yas %>%
  filter(startsWith(hist_tone3, "yang")) %>%
  distri_prop(hist_tone3, sync_tone3, syntax)+
  labs(fill = "synchronic tone")
```
```{r}
unique(f0_tri_ct_yas$syntax)

f0_tri_ct_yas_12mh <- f0_tri_ct_yas %>% filter(syntax == "1+2MH")
f0_tri_ct_yas_21vl <- f0_tri_ct_yas %>% filter(syntax == "2+1VL")
f0_tri_ct_yas_12vo <- f0_tri_ct_yas %>% filter(syntax == "1+2VO")

# 1+2MH
f0_tri_ct_yas_12mh %>% 
 filter(sync_tone1 == 'LHL') %>% 
  distri_prop2(hist_tone23_mapped, sandhi_tone)

# 2+1 VL
f0_tri_ct_yas_21vl %>% 
 filter(sync_tone1 == 'LHL') %>% 
  distri_prop2(hist_tone23_mapped, sandhi_tone)

f0_tri_ct_yas_21vl %>% 
 filter(sync_tone1 == 'LH') %>% 
  distri_prop2(hist_tone23_mapped, sandhi_tone)


# 1+2VO
f0_tri_ct_yas_12vo %>% 
 filter(sync_tone1 == 'LHL') %>% 
  distri_prop2(hist_tone23_mapped, sandhi_tone)

f0_tri_ct_yas_12vo %>% 
 filter(sync_tone1 == 'LH') %>% 
  distri_prop2(hist_tone23_mapped, sandhi_tone)
```